---
title: "p8105_hw2_sk5152"
author: "Shuting Kang"
date: "2022-10-03"
output: github_document
---

## Problem 1
####1. read and clean the data (convert the entry variable from character to logical variable)
```{r}
library(tidyverse)
getwd()
NYC_Subway=read_csv(file='./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv')

# clean up variable names after importing data
NYC_Subway=janitor::clean_names(NYC_Subway)


```
## Problem 2 
#### Read and clean the Mr.Trash Wheel datasets
```{r}

MrTrashWheel=
  readxl::read_excel("./data/Trash Wheel Collection Data.xlsx", sheet="Mr. Trash Wheel") %>%
  janitor::clean_names()%>%
  mutate(dumpster=as.numeric(dumpster))%>%
  filter(!is.na(dumpster))%>%
  mutate(sports_balls=round(sports_balls,0))%>%
  mutate(sports_balls=as.integer(sports_balls))%>%
  select(-x15,-x16)
MrTrashWheel
```
#### import, clean, and organize the data for Professor Trash Wheel
```{r}
ProfTrashWheel=
  readxl::read_excel("./data/Trash Wheel Collection Data.xlsx", sheet="Professor Trash Wheel")%>%
  janitor::clean_names()%>%
  mutate(dumpster=as.numeric(dumpster))%>%
  add_column(sports_balls = NA)%>%
  filter(!is.na(dumpster))
ProfTrashWheel
```
#### combine Professor Trash Wheel with Mr. Trash Wheel
```{r}
MTW=mutate(MrTrashWheel,resource='M')
MTW
PTW=mutate(ProfTrashWheel,resource='P')
TW=rbind(MTW,PTW)
TW
```
#### dataset description 
The combined DataSets contained `r nrow(TW)` observations 
There were `r ncol(TW)` variables, included dumpster,month, year, date,the weight of trash in tons, the number of glass bottles, etc.
During combination procedure, additional character variable 'resource' was added to keep track of which trash wheel is which.
The total weight of trash collected by Professor Trash Wheel was 
`r sum(filter(TW,resource=='P')$weight_tons)` tons;
The total number of sports balls collected by Mr.Trash Wheel was
`r sum(filter(TW,resource=='M',year==2020)$sports_balls)`in 2020.

##Problem 3

#### clean and data in pols-month
```{r}
pols_month=
  read_csv("./data/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(mon,c("year","month","day"))%>%
  mutate(month=month.abb[as.numeric(month)])%>%
  mutate(president=ifelse(prez_gop==1,'gop','dem'))%>%
  select(-prez_dem,-prez_gop,-day)%>%
  mutate(year=as.integer(year))
pols_month
```
#### clean the data in snp.csv
```{r}
snp=
  read_csv("./data/snp.csv")%>%
  janitor::clean_names()%>%
  separate(date,c("month","day","year"))%>%
  mutate(month=month.abb[as.numeric(month)],
         year=as.integer(year),
         year=as.integer(year))%>%
  mutate(year=case_when(year<22 ~year+2000,
                        year>22 ~year+1900))%>%
  select(-day, year,month,everything())
snp
```
#### tidy the unemployment data so that it can be merged with the previous datasets
```{r}
unemployment=
  read_csv("./data/unemployment.csv")%>%
  janitor::clean_names()%>%
  pivot_longer(jan:dec,names_to = "month",values_to = "unemployment")%>%
  mutate(month=str_to_title(month),year=as.integer(year))
unemployment
```
#### join the datasets by merging snp into pols, and merging unemployment into the result
```{r}
snp_to_pols=left_join(pols_month,snp,by=c("year","month"))
snp_to_pols
SPU=left_join(snp_to_pols,unemployment,by=c("year","month"))
SPU
```
#### write a short paragraph about these datasets, describe the result of the dataset
##### e.g. give the dimension, range of years, and names of key variables 

the file "SPU" (pols-month+snp+unemployment datasets) contains `r nrow(SPU)` observations of `r ncol(SPU)` variables related to the number of national politicians who are democratic or republican at any given time, Standard&Poor's stock market index at given time, and percentage of unemployment at given time. The time composed by year and month. 

The key variables include the year and month of the observation and count,  the number of republican/democratic governors on the associated date, the number of republican/democratic senators on the associated date, indicator of whether the president was repulican/democratic on the associated date,the closing values of the S&P stock index on the associated date, and percentage of unemployment in each month of the associated year. 

The range of years for file "SPU" across from `r min(SPU$year)` to `r max(SPU$year)`
However,the record for snp and unemployment at some given time were missing.
